# Basic

In this book, we mainly focus on ggplot2 for generating plots.

Here, we briefly introduce the basic grammar of ggplot2.

> This chapter is mainly adapted from the book *R for Data Science* (<https://r4ds.had.co.nz/data-visualisation.html>).

## Prerequisites

Install the ggplot2 and tidyverse package

> patchwork is a very useful package for combining multiple plots

```{r, eval=FALSE}
install.packages("ggplot2")
install.packages("tidyverse")
install.packages("patchwork")
```

Load the packages. We use the `mpg` dataset for illustration.

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(patchwork)

data("mpg")
mpg
```

## Creating a ggplot

We use scatter plot (`geom_point`) to plot `displ` on x-axis and `hwy` on y-axis:

```{r}
# scatter plot 
# hwy：highway miles per gallon
# displ：engine displacement, in litres
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

In ggplot2, we use the `data` argument to specify the dataset that we want to plot, and assign variables inside `aes()` function, such as `mapping = aes(x = displ, y = hwy)` .

To create different plot, we use the `geom_*` function, such as scatter plot (`geom_point`) and barplot (`geom_bar`). We will go thorough the basic plot types in following chapters.

## Aesthetic mappings

Beside plotting the data, we can assign additional features to point, such as color, size and shape. In ggplot, those properties are called as **aesthetic**.

*For example, you can map the colors of your points to the `class` variable to reveal the class of each car.*

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Or size of the point

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

You can also modify the color of the geom, this will **override** the color assign in `aes`.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class), color = "blue")
```

> Common aesthetic arguments include:
>
> `color`/`colour` and `fill`: for mapping the colors. To be noted, scatter plot only accept `color/colour` for mapping colors.
>
> `shape`: for mapping the shape of the point.
>
> `alpha`: for mapping the transparency.
>
> `size`: for mapping the size.

## Facets

To split the plot into subplots that each display one subset of the data, is facets.

You can use either `facet_wrap()` or `facet_grid()` for facet.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

or the combination of two variables

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

## Geometric objects

We can also combine different geom in one plot

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

# equivalent with the following codes: 
# ggplot(data = mpg) + 
#   geom_point(mapping = aes(x = displ, y = hwy)) +
#   geom_smooth(mapping = aes(x = displ, y = hwy))
```

we can specify different aesthetic and different data in each geom

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

## Statistical transformation

Next, we use `diamonds` dataset for illustration.

```{r}
data("diamonds")
diamonds
```

ggplot2 contains lots of useful statistical transformation for visualization. For example, if we want to count the number of diamonds grouped by cut, we can simply use barplot to do this job.

```{r}
table(diamonds$cut)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut), stat = "count")

# equals to:
# ggplot(data = diamonds) + 
#   stat_count(mapping = aes(x = cut))
```

In `count` mode, `geom_bar` can group the variables by `x` and count the frequency of each member.

Also, you may be interested with other statistics

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

## Position adjustment

The color of barplot can be assigned with `color` and `fill`

```{r, fig.width=10, fig.height=4}
p1 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
p2 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))

# patchwork for combining plots
p1 + p2
```

Beside, color can be grouping variable for displaying different property.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position='identity')
```

Each layer of the barplot represents the number of diamonds with different clarity.

`position` have three options: `"identity"`, `"dodge"` or `"fill"`.

-   `position = "fill"`

    Show the proportions of each group

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

-   `position = "dodge"`

    Place overlapping objects beside one another

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

## Coordinate systems

### coord_flip()

`coord_flip()` switches the x and y axes.

```{r, fig.width=8, fig.height=3}
p1 <- ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
p2 <- ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
p1 + p2
```

### coord_polar()

`coord_polar()` uses polar coordinates.

```{r}
ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL) +
  coord_polar()
```

## Summary

In all, ggplot2 can be summarized as the following code template:

```{r, eval=FALSE}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```
