# Barplot

In this chapter, we introduce the barplot in ggplot2 style.

## basic barplot 

We use `diamonds` dataset for illustration.

```{r}
library(ggplot2)
library(dplyr)
library(patchwork)

data("diamonds")
diamonds
```

Count the number of diamonds in each cut group.

```{r}
df1 <- diamonds %>% 
  count(cut)
df1
```

We use `stat="identity"` mode to generate barplot

```{r}
ggplot(df1, aes(x=cut, y=n)) +
  geom_bar(stat="identity")
```


## With color

```{r}
ggplot(df1, aes(x=cut, y=n)) +
  geom_bar(aes(fill=cut), stat="identity")
```

Using `fill` as grouping variable

```{r}
df2 <- diamonds %>% 
  count(cut, clarity)
df2
```

```{r}
ggplot(df2, aes(x=cut, y=n)) +
  geom_bar(aes(fill=clarity), stat="identity", position="identity")
```

### adjust position

There are two other more position options: "fill" and "dodge"

```{r, fig.width=6, fig.height=4}
p1 <- ggplot(df2, aes(x=cut, y=n)) +
  geom_bar(aes(fill=clarity), stat="identity", position="fill")

p2 <- ggplot(df2, aes(x=cut, y=n)) +
  geom_bar(aes(fill=clarity), stat="identity", position="dodge")

p1 + p2
```

## Horizontal barplot

```{r}
ggplot(df1, aes(x=cut, y=n)) +
  geom_bar(aes(fill=cut), stat="identity") +
  coord_flip()
```

## Circular barplot


```{r}
ggplot(df1, aes(x=cut, y=n)) +
  geom_bar(aes(fill=cut), stat="identity", width=1) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL) +
  coord_polar()
```