# Heatmap

In this chapter, we introduce the heatmap in ggplot2, pheatmap and ComplexHeatmap styles. 

Install related packages 

```{r, eval=FALSE}
install.packages(c("ggplot2","pheatmap","ComplexHeatmap"))
```

Load packages and create dummy data

```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(pheatmap)
library(ComplexHeatmap)
library(paintingr) # for generation of color palettes

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
dum.data <- expand.grid(X=x, Y=y)
dum.data$Z <- seq(1,20)+runif(400, 0, 5)

head(dum.data)
```

## ggplot2 

Basic heatmap in ggplot2 style (`geom_tile`)

```{r}
# Heatmap 
pal <- paint_palette("Autumn", n=100, type="continuous")
ggplot(dum.data, aes(X, Y, fill= Z)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal) + 
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) + 
  coord_equal() 
```

## pheatmap

Both pheatmap and ComplexHeatmap require matrix like input

```{r}
# Dummy data
set.seed(123)
dum.mat <- matrix(rnorm(1000), ncol=10)

head(dum.mat)
```

Create basic pheatmap

```{r}
pheatmap::pheatmap(dum.mat)
```

### Control color

```{r}
pal2 <- paint_palette("Twilight", 100, "continuous")
pheatmap::pheatmap(dum.mat, color = pal2, border_color = NA)
```

## ComplexHeatmap

Create basic ComplexHeatmap

```{r}
ComplexHeatmap::Heatmap(dum.mat, name="mat")
```

### Control color

Pass vector of colors to change the color

```{r}
ComplexHeatmap::Heatmap(dum.mat, name="mat", col=paint_palette("Vesuvius"))
```

Color of ComplexHeatmap can also be modified with a color mapping function.

```{r}
# generate color mapping function
col_fun <- circlize::colorRamp2(breaks = seq(-3,3), 
                                colors = paint_palette("Twilight",7,"continuous"))
ComplexHeatmap::Heatmap(dum.mat, 
                        name="mat", 
                        col=col_fun, 
                        heatmap_legend_param = list(at = c(-3,-1.5,0,1.5,3)) # control breaks of heatmap legend
                        )
```